### 2013-02-15 patchy-selection-paper.tex
/* Patchy Selection paper */

%%%%%%%%%%
\section{Applications} 
Hoekstra, Nachman, and colleagues have made an extensive study of coat color in the rock pocket mouse, \emph{Chaetodipus intermedius}, building on classic work by XXX. These mice live in the desert? through the American SouthWest and into Mexico. In the majority of their range they live on the light colored rocks, and have light pigmentation \citep[presumably an adaptation to visual-based predators][]{}. However, in some regions these mice live on dark colored substrates, including old lava flows, and have adapted to these patches in part through changes to a dark pigmentation. On one of these flows (Pinacate) an allele at the Mc1r locus is responsible for much of the change to a dark pelage \citep{Nachman:03}. This dark allele differs from the light allele by four amino acid changes - and has a dominant or partially dominant effect depending on the measure of coat color assessed. These lava flows are separated from each other by a range of distances, with some being separated from any other flow by hundreds of kilometers of light colored substrate. The Pinacate allele is not present in a number of other dark colored rock pocket mouse populations suggesting these populations have adapted in parallel. However, there is some evidence \citep{Hoekstra:05} that some close populations may have adapted by the spread of another dark allele by migration.

As can be seen from above a key parameter is $\sigma/\sqrt{s}$. \citep{Hoekstra:04} studied the frequency of the dark MC1R allele, and coat color phenotypes, at sites on the Pinacate lava flow and at two nearby light substrate sites. On the lava flow the dark MC1R allele is at $86\%$ frequency. At the Tule light substrate site, which is $\sim 12km$ from the flow, the dark MC1R allele was at $\sim 3\%$ frequency. At a greater distance from the lava ($XXX$km) the light colored substrate the dark MC1R was absent from a sample at the Christmas pass. While at the light substrate O'Neill site, which is $10$km from the flow on opposite side from Tule, the dark MC1R allele has a frequency of $\sim 34\%$.

I estimate that a characteristic length of 2.8km gives the frequency dropping from 50\% to $\sim$1\% in this distance. So this doesn't seem so off. A characteristic length that is 1.4x gives 2.8\% from 50\% in 12km, which may be a little more accurate.

2) The frequency at O'Neill [another site off the lava, on opposite side from Tule]  is 9km from the lava and has a frequency of 34\% of the dark allele. I estimate that we need a sigma/sqrt(s) of $\sim$28km (i.e.\ ten times the characteristic length) to give a drop from 50\% to 34\%.

[Obviously the cline isn't decaying from 50\%, the frequency of the light allele on the lava to be $\sim$14\%, we could put this in more carefully]
So it seems to me that a sigma/sqrt(s) of 3-30km seems to encompass a range of interesting parameters, and can be justified without too much trouble.

Using the dispersal distance of 0.28km [from marker recapture] and an s=0.01 [from paper] gives us sigma/sqrt(s)=2.8km

HbS

G6PD

>>>>

%%%%%%%%%%
\section{Applications} 
Hoekstra, Nachman, and colleagues have made an extensive study of coat color in the rock pocket mouse, \emph{Chaetodipus intermedius}, building on classic work by XXX. These mice live throughout the American SouthWest and into Mexico, and mostly live on the light colored rock outcrops, and have light pigmentation \citep[presumably an adaptation to visual-based predators][]{}. However, in some regions these mice live on dark colored substrates, including old lava flows, and have adapted to these patches in part through changes to a dark pigmentation. On one of these flows (Pinacate) an allele at the Mc1r locus is responsible for much of the change to a dark pelage \citep{Nachman:03}. This dark allele differs from the light allele by four amino acid changes - and has a dominant or partially dominant effect depending on the measure of coat color assessed. These lava flows are separated from each other by a range of distances, with some being separated from any other flow by hundreds of kilometers of light colored substrate. The Pinacate allele is not present in a number of other dark colored rock pocket mouse populations suggesting these populations have adapted in parallel. However, there is some evidence \citep{Hoekstra:05} that, elsewhere in the range, some nearby populations may share a dark phenotype whose genetic basis has been spread by migration.

As can be seen from above a key parameter is the characteristic lengths $\sigma/\sqrt{s_m}$. \citep{Hoekstra:04} studied the frequency of the dark MC1R allele, and coat color phenotypes, at sites on the Pinacate lava flow and at two nearby light substrate sites.
 On the lava flow the dark MC1R allele is at $86\%$ frequency. At the Tule light substrate site, which is $\sim 12km$ from the flow, the dark MC1R allele was at $\sim 3\%$ frequency. 
At a greater distance from the lava ($XXX$km) the light colored substrate the dark MC1R was absent from a sample at the Christmas pass. 
While at the other light substrate site (O'Neill), which is $10$km from the flow on opposite side from Tule, the dark MC1R allele has a frequency of $\sim 34\%$. On the basis of these numbers we can gain a sense of a plausible range of characteristic lengths.
Assuming the allele is at $50\%$ frequency at the edge of the lava flow, 
we can fit the frequency found at each of the light substrate sites that are 
polymorphic for the dark allele under a simple allele frequency cline model \citep{XXX}.
Doing this for the Tule we obtain $\sigma/\sqrt{s_m} \approx 4$km and for the O'Neill site
$\sigma/\sqrt{s_m} \approx 30$km. \\

The mutational target size $\mu$ for the trait is unclear. 
While the Pinacate dark haplotype differs from the light haplotype at three amino-acid residues,
it is likely that not all of these changes are needed for a population to begin to  
adapt. Also there a number of genes, beyond MC1R, at which adaptative changes affecting 
pigmentation have been identified in closely related species and more
broadly across vertibrates \citep{}.To span a range of plausible
values, we use a low mutation rate 
We use a low mutation rate $\mu= 10^{-8}$, where the adaptive
mtation has to occur at a single base pair.
An intermediate mutation rate $\mu= 10^{-5}$, equivalent to
the mutation being able to arise at any of a thousand bases pairs,
e.g. any of the coding bases of a typical length gene. And a high
mutation rate $\mu = 10^{3}$, corresponding to the adaptive mutation
being able to arise in any of $100$ genes, e.g. anywhere in a large pathway.
